---
title: "Age and Condition Differences in Alpha Desynchronization Across Infancy: Cross-sectional and Longitudinal Study"
author: "Mina Elhamiasl"
date:  "December 2023"
output: html_document
---
**======================================================================================================================**
**"After data pre-processing and computing desynchronization values using MATLAB, calculate the average desynchronization within the desired frequency range (here 6 to 9Hz) for each electrode and participant. Save the electrode*participant data in a long format .csv file for both the cross-sectional and longitudinal samples**
**Then, use this code to analyze (ANOVA and T-test) and visualize both cross-sectional and longitudinal data**
**======================================================================================================================**
`

Install all required packeages.
```{r install packages}
packages <- c("cowplot", "readr", "ggplot2", "dplyr", "lavaan", "smooth", "Hmisc", "tidyverse", "reshape2", "rio", "Rmisc", "purrr", "magrittr", "yarrr","tidyr", "forcats", "esquisse" , "yarrr", "broom", "purrr", "FSA", "rstatix", "afex")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))
}
```

```{r libraries}
library(tidyverse)
library(dplyr)
library(rio)
library(reshape2)
library(ggplot2)
library(Rmisc)
library(purrr)
library(magrittr)
library(cowplot)
library(readr)
library(lavaan)
library(smooth)
library(Hmisc)
library(tidyr)
library(yarrr)
library(forcats)
library(esquisse)
library(yarrr)
library(broom)
library(purrr)
library(FSA)
library(rstatix)
library(afex)
```


**======================================================================================**
DATA: CROSS-SECTIONAL
AGE GROUPS: 6, 9, and 12 months old
CONDITIONS: Faces & Objects
ROI: OCCIPITAL ELECTRODES
**======================================================================================**
*=======================================================================================*
*1. Data Preparation*
*=======================================================================================*

Set the directory
```{r}
setwd("/Users/mina/OneDrive - University of Florida/one drive/Alpha_desynch_2023")
```

Imported the data in .csv format
```{r Raw data}
raw<-read.csv('/Users/mina/OneDrive - University of Florida/one drive/Alpha_desynch_2023/alpha_desynch_cross.csv')
names(raw)
```


If you have any missing data and you prefer to remove them. If you do not have any, skip this section
```{r}
no_missing <- na.omit(raw)                           # Apply na.omit function
no_missing
```


Convert a column from numeric to factor if needed.
```{r}
raw$id<-as.factor(raw$id)
raw$age<-as.factor(raw$age)
raw$condition<-as.factor(raw$condition)
```

Select the required variables.
```{r}
main_cross_occipital<-select(raw, c("id", "age", "condition", "E70", "E71", "E74","E75", "E76", "E82", "E83"))
names(main_cross_occipital)
```

Calculate the average of alpha desycnhronization across occipital region. 
```{r}
#main_cross_occipital <- main_cross_occipital %>% 
  #mutate(desynch_occipital = rowMeans(select(., c("E70", "E71", "E74", "E75", "E76", "E82", "E83"))))

#OR

main_cross_occipital$desynch_occipital <- with(main_cross_occipital, ((E70+E71+E74+E75+E76+E82+E83)/7))
main_cross_occipital
```

Select the required variables.
```{r}
main_cross_occipital<-select(main_cross_occipital, c("id", "age", "condition", "desynch_occipital"))
names(main_cross_occipital)
```

Change the data to wide format. 
```{r}
data_wide_cross_occipital<-main_cross_occipital%>%
  tidyr::spread(condition, desynch_occipital)
#write.csv(data_wide_cross_occipital,"data_wide_cross_occipital.csv")
data_wide_cross_occipital
```

Change the data to long format for plotting and further analysis.
```{r}
data_long_occipital_cross<-gather(data_wide_cross_occipital,  condition,amplitude, face:obj)
data_long_occipital_cross
```

Check the outliers if needed
```{r}
data_long_occipital_cross %>%
  group_by(age, condition) %>%
  identify_outliers(amplitude)
```

Remove the outlier if needed
```{r}
#data_long_occipital_cross_Nooutlier <- subset(data_long_occipital_cross, id != "202")
``` 

*=======================================================================================*
*2. Data Analysis*
*=======================================================================================*

Descriptive information (desynchronization amlitudes) per age, condition and age*condition
```{r descriptive for condition}
o_Sum = Summarize(amplitude ~ condition,
                data=data_long_occipital_cross,
                digits=3)
o_Sum$se = o_Sum$sd / sqrt(o_Sum$n)
o_Sum$se = signif(o_Sum$se, digits=3)
o_Sum
```

```{r descriptive for age}
o_Sum = Summarize(amplitude ~ age,
                data=data_long_occipital_cross,
                digits=3)
o_Sum$se = o_Sum$sd / sqrt(o_Sum$n)
o_Sum$se = signif(o_Sum$se, digits=3)
o_Sum
```

```{r descriptive for age*condition}
o_Sum = Summarize(amplitude ~ condition+age,
                data=data_long_occipital_cross,
                digits=3)
o_Sum$se = o_Sum$sd / sqrt(o_Sum$n)
o_Sum$se = signif(o_Sum$se, digits=3)
o_Sum
```

Mixed ANOVA analysis using oav_ez function to examine the main effects of age, condition, and their interaction
```{r}
mixed_anova <- aov_ez(
  id = "id",  # participants
  dv = "amplitude", # dependent variable
  data = data_long_occipital_cross,
  between = "age", # between-group factor
  within = c ("condition") # within-group factor
)
get_anova_table(mixed_anova)
```

Folow up the main effects using paired or independent t-tests.
Firts, change the data to wide format.
```{r}
data_wide_occipital_cross_t<-data_long_occipital_cross%>%
  tidyr::spread(condition, amplitude)
write.csv(data_wide_occipital_cross_t,"data_wide_occipital_cross_t.csv")
data_wide_occipital_cross_t
```

Calculate the average of condition
```{r}
data_wide_occipital_cross_t$mean_cond_t <- with(data_wide_occipital_cross_t, ((face+obj)/2)) 
data_wide_occipital_cross_t
```

Chnage it to the long format
```{r}
data_long_mean_occipital_cross<-gather(data_wide_occipital_cross_t,  condition,amplitude, face:mean_cond_t)
data_long_mean_occipital_cross
```

Subset the data for each of 6-, 9, and 12-months-old groups for the t-test analysis .
```{r}
t_6mo <- subset(data_long_mean_occipital_cross,  age == "6"&condition == "mean_cond_t", amplitude,
                 drop = TRUE)
t_9mo <- subset(data_long_mean_occipital_cross,  age == "9"&condition == "mean_cond_t", amplitude,
                 drop = TRUE)
t_12mo <- subset(data_long_mean_occipital_cross,  age == "12"&condition == "mean_cond_t", amplitude,
                 drop = TRUE)

```

T-Test analysis 
```{r difference between 6mo and 9mo}
t.test(t_6mo, t_9mo, paired = FALSE, alternative = "two.sided")
```

```{r difference between 6mo and 12mo}
t.test(t_6mo, t_12mo, paired = FALSE, alternative = "two.sided")
```

```{r difference between 9mo and 12mo}
t.test(t_9mo, t_12mo, paired = FALSE, alternative = "two.sided")
```

Subset the data for each age group (6-, 9, and 12-months-old) per  condition (Face and Object).
```{r}
t_6mo_fa <- subset(long_mean,  age == "6"&condition == "face", amplitude,
                 drop = TRUE)
t_9mo_fa <- subset(long_mean,  age == "9"&condition == "face", amplitude,
                 drop = TRUE)
t_12mo_fa <- subset(long_mean,  age == "12"&condition == "face", amplitude,
                 drop = TRUE)

t_6mo_obj <- subset(long_mean,  age == "6"&condition == "obj", amplitude,
                 drop = TRUE)
t_9mo_obj <- subset(long_mean,  age == "9"&condition == "obj", amplitude,
                 drop = TRUE)
t_12mo_obj <- subset(long_mean,  age == "12"&condition == "obj", amplitude,
                 drop = TRUE)

```

T-Test
```{r between age differences in face cobdition}
t.test(t_6mo_fa, t_9mo_fa, paired = FALSE, alternative = "two.sided")
t.test(t_6mo_fa, t_12mo_fa, paired = FALSE, alternative = "two.sided")
t.test(t_9mo_fa, t_12mo_fa, paired = FALSE, alternative = "two.sided")
```

```{r between age differences in object cobdition}
t.test(t_6mo_obj, t_9mo_obj, paired = FALSE, alternative = "two.sided")
t.test(t_6mo_obj, t_12mo_obj, paired = FALSE, alternative = "two.sided")
t.test(t_9mo_obj, t_12mo_obj, paired = FALSE, alternative = "two.sided")
```

*=======================================================================================*
*3. Data Visualization*
*=======================================================================================*
*Visualize the main effects of age and condition as well as their interactions*

Interaction of age and condition. 
```{r}
figure_age<-pirateplot(amplitude ~ age+condition ,
           data=data_long_occipital_cross,
           theme = 1,
           main = "",
           #xlab = "age",
           ylab = "desynchronization",
           pal = "info",  #color pallate
           #ylim=c(1500,2700),
           point.pch = 19,
           back.col = transparent("white", .1), #add light purple background
           inf.f.o = .0, # Turn off inf fill
           inf.b.o = .9, # Turn off inf border
           point.o = .9,   # Turn up points
           bar.f.o = .9, # Turn up bars
           bean.f.o = .5, # Light bean filling
           bean.b.o = .9, # Light bean border
           avg.line.o = .9, # Average line
           inf.b.col = "black", # Inf border col
           avg.line.col = "black") # avg line colrea
```

For the main effect of age, aggregate the data across condition and select the requeired variables. 
```{r}
data_long_occipital_cross_id_age = aggregate(data_long_occipital_cross$amplitude, list( data_long_occipital_cross$id, data_long_occipital_cross$age), mean)
colnames(data_long_occipital_cross_id_age) = c("id", "age", "amplitude")
```

```{r}
figure_age<-pirateplot(amplitude ~ age ,
           data=data_long_occipital_cross_id_age,
           theme = 1,
           main = "",
           #xlab = "age",
           ylab = "desynchronization",
           pal = "info",  #color pallate
           #ylim=c(1500,2700),
           point.pch = 19,
           back.col = transparent("white", .1), #add light purple background
           inf.f.o = .0, # Turn off inf fill
           inf.b.o = .9, # Turn off inf border
           point.o = .9,   # Turn up points
           bar.f.o = .9, # Turn up bars
           bean.f.o = .5, # Light bean filling
           bean.b.o = .9, # Light bean border
           avg.line.o = .9, # Average line
           inf.b.col = "black", # Inf border col
           avg.line.col = "black") # avg line colrea
```

For the main effect of condition aggregate the data across ages, and select the requeired variables. 
```{r}
data_long_occipital_cross_id_condition = aggregate(data_long_occipital_cross$amplitude, list( data_long_occipital_cross$id, data_long_occipital_cross$condition), mean)
colnames(data_long_occipital_cross_id_condition) = c("id", "condition", "amplitude")
```

```{r}
figure_age<-pirateplot(amplitude ~ condition ,
           data=data_long_occipital_cross_id_condition,
           theme = 1,
           main = "",
           #xlab = "age",
           ylab = "desynchronization",
           pal = "pony",  #color pallate
           #ylim=c(1500,2700),
           point.pch = 19,
           back.col = transparent("white", .1), #add light purple background
           inf.f.o = .0, # Turn off inf fill
           inf.b.o = .9, # Turn off inf border
           point.o = .9,   # Turn up points
           bar.f.o = .9, # Turn up bars
           bean.f.o = .5, # Light bean filling
           bean.b.o = .9, # Light bean border
           avg.line.o = .9, # Average line
           inf.b.col = "black", # Inf border col
           avg.line.col = "black") # avg line colrea
```


**======================================================================================**
DATA: LONGITUDINAL
AGE GROUPS: 6, 9, and 12 months old
CONDITIONS: Faces & Objects
ROI: OCCIPITAL ELECTRODES
**======================================================================================**
*=======================================================================================*
*1. data preparation*
*=======================================================================================*

Set the directory
```{r}
setwd("/Users/mina/OneDrive - University of Florida/one drive/Alpha_desynch_2023")
```

Imported the data in .csv format
```{r Raw data}
raw<-read.csv('/Users/mina/OneDrive - University of Florida/one drive/Alpha_desynch_2023/alpha_desynch_Longitudinal.csv')
names(raw)
```

Convert a column from numeric to factor if needed.
```{r}
raw$id<-as.factor(raw$id)
raw$age<-as.factor(raw$age)
raw$condition<-as.factor(raw$condition)
```

Select the required variables.
```{r}
main_Longitudinal_occipital<-select(raw, c("id", "age", "condition", "E70", "E71", "E74","E75", "E76", "E82", "E83"))
names(main_Longitudinal_occipital)
```

Calculate the average of alpha aLongitudinal occipital region. 
```{r}
#main_Longitudinal_occipital <- main_Longitudinal_occipital %>% 
#mutate(desynch_occipital = rowMeans(select(., c("E70", "E71", "E74", "E75", "E76", "E82", "E83"))))

#OR

main_Longitudinal_occipital$desynch_occipital <- with(main_Longitudinal_occipital, ((E70+E71+E74+E75+E76+E82+E83)/7))
main_Longitudinal_occipital
```

Select the required variables.
```{r}
main_Longitudinal_occipital<-select(main_Longitudinal_occipital, c("id", "age", "condition", "desynch_occipital"))
names(main_Longitudinal_occipital)
```

Change the data to wide format and save it just for the ease of use.
```{r}
data_wide_Longitudinal_occipital<-main_Longitudinal_occipital%>%
  tidyr::spread(condition, desynch_occipital)
#write.csv(data_wide_Longitudinal_occipital,"data_wide_Longitudinal_occipital.csv")
data_wide_Longitudinal_occipital
```

Change the data to long format for plotting and further analysis.
```{r}
data_long_occipital_Longitudinal<-gather(occipital_Longitudinal,  condition,amplitude, face:obj)
data_long_occipital_Longitudinal
```

Check the outliers if needed
```{r}
data_long_occipital_Longitudinal %>%
  group_by(age, condition) %>%
  identify_outliers(amplitude)
```

Remove the outlier if needed
```{r}
#data_long_occipital_Longitudinal_Nooutlier <- subset(data_long_occipital_Longitudinal, id != "202")
``` 

*=======================================================================================*
*2. Data Analysis*
*=======================================================================================*

Descriptive information (desynchronization amlitudes) per age, condition and age*condition
```{r descriptive for condition}
o_Sum = Summarize(amplitude ~ condition,
                data=data_long_occipital_Longitudinal,
                digits=3)
o_Sum$se = o_Sum$sd / sqrt(o_Sum$n)
o_Sum$se = signif(o_Sum$se, digits=3)
o_Sum
```

```{r descriptive for age}
o_Sum = Summarize(amplitude ~ age,
                data=data_long_occipital_Longitudinal,
                digits=3)
o_Sum$se = o_Sum$sd / sqrt(o_Sum$n)
o_Sum$se = signif(o_Sum$se, digits=3)
o_Sum
```

```{r descriptive for age*condition}
o_Sum = Summarize(amplitude ~ condition+age,
                data=data_long_occipital_Longitudinal,
                digits=3)
o_Sum$se = o_Sum$sd / sqrt(o_Sum$n)
o_Sum$se = signif(o_Sum$se, digits=3)
o_Sum
```

Repeated measure ANOVA to examine main effects of age and condition and their interaction
```{r}
res.aov <- anova_test(
  data = data_long_occipital_Longitudinal, dv = amplitude, wid = id,
  within = c(age, condition) #within group factors
  )
get_anova_table(res.aov)
```

*======================================================================================*
*3. Data Visualization*
*======================================================================================*
*Visualize the main effects of age and condition as well as their interactions*

Interaction of age and condition
```{r}
figure_age<-pirateplot(amplitude ~ age+condition ,
           data=data_long_occipital_Longitudinal,
           theme = 1,
           main = "",
           #xlab = "age",
           ylab = "desynchronization",
           pal = "info",  #color pallate
           #ylim=c(1500,2700),
           point.pch = 19,
           back.col = transparent("white", .1), #add light purple background
           inf.f.o = .0, # Turn off inf fill
           inf.b.o = .9, # Turn off inf border
           point.o = .9,   # Turn up points
           bar.f.o = .9, # Turn up bars
           bean.f.o = .5, # Light bean filling
           bean.b.o = .9, # Light bean border
           avg.line.o = .9, # Average line
           inf.b.col = "black", # Inf border col
           avg.line.col = "black") # avg line colrea
```

For the main effect of age, aggregate the data across condition and select the requeired variables. 
```{r}
data_long_occipital_Longitudinal_id_age = aggregate(data_long_occipital_Longitudinal$amplitude, list( data_long_occipital_Longitudinal$id, data_long_occipital_Longitudinal$age), mean)
colnames(data_long_occipital_Longitudinal_id_age) = c("id", "age", "amplitude")
```

```{r}
figure_age<-pirateplot(amplitude ~ age ,
           data=data_long_occipital_Longitudinal_id_age,
           theme = 1,
           main = "",
           #xlab = "age",
           ylab = "desynchronization",
           pal = "info",  #color pallate
           #ylim=c(1500,2700),
           point.pch = 19,
           back.col = transparent("white", .1), #add light purple background
           inf.f.o = .0, # Turn off inf fill
           inf.b.o = .9, # Turn off inf border
           point.o = .9,   # Turn up points
           bar.f.o = .9, # Turn up bars
           bean.f.o = .5, # Light bean filling
           bean.b.o = .9, # Light bean border
           avg.line.o = .9, # Average line
           inf.b.col = "black", # Inf border col
           avg.line.col = "black") # avg line colrea
```

For the main effect of condition, aggregate the data across age, and select the requeired variables. 
```{r}
data_long_occipital_Longitudinal_id_condition = aggregate(data_long_occipital_Longitudinal$amplitude, list( data_long_occipital_Longitudinal$id, data_long_occipital_Longitudinal$condition), mean)
colnames(data_long_occipital_Longitudinal_id_condition) = c("id", "condition", "amplitude")
```

```{r}
figure_age<-pirateplot(amplitude ~ condition ,
           data=data_long_occipital_Longitudinal_id_condition,
           theme = 1,
           main = "",
           #xlab = "age",
           ylab = "desynchronization",
           pal = "pony",  #color pallate
           #ylim=c(1500,2700),
           point.pch = 19,
           back.col = transparent("white", .1), #add light purple background
           inf.f.o = .0, # Turn off inf fill
           inf.b.o = .9, # Turn off inf border
           point.o = .9,   # Turn up points
           bar.f.o = .9, # Turn up bars
           bean.f.o = .5, # Light bean filling
           bean.b.o = .9, # Light bean border
           avg.line.o = .9, # Average line
           inf.b.col = "black", # Inf border col
           avg.line.col = "black") # avg line colrea
```


